---
title: "Case2"
author: "Murilo Mendel Costa"
date: "08/07/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dslabs)
library(tidyverse)
library(ggplot2)
library(readr)
library(stringr)
library(scales)
library(forecast)

```

```{r load, include=FALSE}
# Importing data to R environment
# Importing data to R environment
cwd <- getwd()
file <- "/data/data_comexstat.csv"
path <- file.path(cwd, file)

bra_df <- as.data.frame(read_csv(path, col_names = TRUE))

#Data Summary
head(bra_df)
summary(bra_df)

# Set of values for some columns
unique(bra_df$product)
unique(bra_df$state) %>% sort(decreasing = FALSE)
unique(bra_df$country) %>% sort(decreasing = FALSE)
unique(bra_df$route)

# Countries Ocurrences
table(bra_df$country)

```


### Brazilian imports/exports;

#### This document contain a report study about Brazilian monthly imports and exports, mantained by the government from 1997 to 2019.
#### The data set is composed by 117965 rows, 8 columns and has no missing value. The 8 columns are describe as:
#### 1. date: date variable type from 1997-01-01 to 2019-12-01
#### 2. product (categorical char variable): represents the product imported/exported: [corn, soybean_meal, soybean_oil, soybeans, sugar, wheat]
#### 3. state (Categorical char feature: Brazilian states abreviation from/to where the export/import are related: all 26 brazilian states initials.
#### 4. country (Categorical char feature): Countries from/to where the export/import are to/from: 212 different countries.
#### 5. type (Categorical char feature): type of commercial operation: [Export, Import]
#### 6. route (categorical char feature): means the transportation way for the commercial operation: [Air, Ground, River, Sea, Other]
#### 7. tons (numerical float feature): total weight quantity exported/imported in tons.
#### 8. usd (numerical integer value): price over the operation in USD.

### Evolution of total monthly exports from Brazil of Soybeans, Soybean Oil and Soybean Meal

```{r pre_bsme, include=FALSE}
separate_date <- bra_df %>% separate(date, c("year", "month", "day"), "-")
bra_df <- bra_df %>% mutate(year = separate_date$year)
bra_df[,9] <- sapply(bra_df[,9], as.numeric)
head(bra_df)
```

```{r brazilian_soybean_monthly_report, include=FALSE}
# Creating a monthly summary oin tons for each product
bra_soy_export_monthly <- bra_df %>% filter(type == "Export") %>%
    filter(product == c("soybean_meal", "soybean_oil", "soybeans")) %>%
    group_by(date, product) %>%
    summarise(tons = sum(tons))
```

```{r month_plot, echo=FALSE, fig.width=14, fig.height=6}
#Plotting the monthly Summary of brazilian exported soybeans
bra_soy_export_monthly %>% ggplot() +
  geom_line(aes(x = date, y = tons, color = product)) +
  scale_y_continuous(trans = 'log10', 
                     labels = trans_format("log10", math_format(10^.x)), 
                     breaks = trans_breaks("log10", function(x) 10^x)) +
  scale_x_date(breaks = date_breaks("years"), 
               labels = date_format("%m/%y")) +
  theme(axis.text.x = element_text(angle = 80),
        plot.title = element_text(hjust = 0.5, size = 16)) +
  labs(title = "Monthly Brazilian Soybeans Export",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  xlab("Months)") +
  ylab("Quantity(tons)")
```

#### It's possible to recognize some general patterns graphically:<br/>
#### - The average monthly export of Soybeans is the biggest, however it has the biggest variance too.
#### - The average monthly export of Soybean Meal is the second biggest and has a lower variance than Soybeans.
#### - The average monthly export of Soybean Oil is the smallest of these three products.

```{r brazilian_soybean_yearly_export, include=FALSE}
# Creating a yearly export summary in tons for each product
bra_soy_export_yearly <- bra_df %>% filter(type == "Export") %>%
  filter(product == c("soybean_meal", "soybean_oil", "soybeans")) %>%
  group_by(year, product) %>%
  summarise(tons = sum(tons))
```

```{r year_plot, echo=FALSE, fig.width=14, fig.height=6}
#Plotting the yearly Summary of brazilian exported soybeans
bra_soy_export_yearly %>% ggplot() +
  geom_line(aes(x = year, y = tons, color = product), lwd = 1.1) +
  scale_y_continuous(trans = log10_trans(),
                     limits = c(100000, 100000000),
                     labels = trans_format("log10", math_format(10^.x)), 
                     breaks = c(100000,1000000, 10000000, 100000000)) +
  scale_x_continuous(limits = c(1997,2019),
                     breaks = seq(1997,2019,1)) +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(hjust = 0.5, size = 16)) +
  labs(title = "Yearly Brazilian Soybeans Export",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  xlab("Years") +
  ylab("Quantity(tons)")
```

#### From the Yearly report:
#### - Soybeans are the main exported product between the three observed, by soybean meal and soybean Oil.
#### - Soybeans and Soybean meal export quantity increased over time.

## Top 3 most important products exported in the last 5 years

```{r rank_5year_export, include=FALSE}
# Summary of every export in the last 5 years
rank_export_5year <- bra_df %>% filter(type == "Export" & year >= 2015) %>%
  group_by(product) %>%
  summarise(total_tons = sum(tons), total_USD = sum(usd), yearly_profit = mean(total_USD/5)) %>%
  arrange(desc(yearly_profit))

rank_export_5year <- data.frame(rank_export_5year)
```

```{r show_rank_table, echo=FALSE}
print(rank_export_5year)
```

```{r 5year_rank_plot, echo=FALSE, fig.width=14, fig.height=8}
## Plotting the last 5 year export results from Brazil from the economic point of view
rank_export_5year %>% ggplot() +
  geom_col(aes(x = reorder(product, -yearly_profit), y = yearly_profit, fill = reorder(product, -yearly_profit))) +
  scale_y_continuous(label = dollar,) +
  labs(title = "Ranking about brazilian exports in the last 5 years",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  theme(axis.text.x = element_text(angle = 45 ),
        plot.title = element_text(hjust = 0.5, size = 16),
        legend.title = element_blank()) +
  xlab("Products") +
  ylab("Profit(USD/year)")
```

#### The 3 most important products exported by Brazil in the last five years was:
#### 1. Soybeans: 24,665,105,971 [$/year].
#### 2. Sugar: 8,187,697,204 [$/year]
#### 3. Soybean Meal: 5,681,544,687 [$/year]

### Routes Study:

#### The main route through corns are exported from Brazil are from Sea, due to the following table:

```{r corn_route_summary, include=FALSE}
corn_export_route <- bra_df %>% filter(product == "corn" & type == "Export" & year >= 2015) %>%
  group_by(route) %>%
  summarise(total_tons = sum(tons)) %>%
  arrange(desc(total_tons))

corn_export_route <- data.frame(corn_export_route)

product_vs_route <- bra_df %>% filter(type == "Export" & year >= 2015) %>%
  group_by(product, route) %>%
  summarise(total_tons = sum(tons)) %>%
  arrange(product, desc(total_tons))
```

```{r show_corn_route_table, echo=FALSE}
print(corn_export_route)
```

#### The next analysis shows that there is not a difference between products and routes for exportats in Brazil, since the main export route are by Sea for every product.

```{r export_route_plot, echo=FALSE, fig.width=14, fig.height=8}
product_vs_route %>% ggplot() +
  geom_bar(aes(x = product, y = total_tons, fill = reorder(route, -total_tons)), 
           position = "dodge", stat = "identity") +
  scale_y_continuous(trans = log10_trans(),
                     labels = trans_format("log10", math_format(10^.x))) +
  labs(title = "Main export routes from Brazil",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  theme(axis.text.x = element_text(angle = 45 ),
        plot.title = element_text(hjust = 0.5, size = 16),
        legend.title = element_blank()) +
  xlab("Products") +
  ylab("Quantity(tons)")
```

### Trade Partners:
#### The following analysis shows the top 10 most important trade partners for Brazil considering corn and sugar commercial trades over the last 3 years.

```{r corn_trade_partner_summary, include=FALSE}
# Most important trade partners for brazil for corn in the last 3 years
# Import
trade_partner_corn_import <- bra_df %>% filter(product == "corn" & year <= 2017 & type == "Import") %>%
  group_by(country) %>%
  summarise(total_import = sum(tons)) %>%
  arrange(desc(total_import))

# Export
trade_partner_corn_export <- bra_df %>% filter(product == "corn" & year <= 2017 & type == "Export") %>%
  group_by(country) %>%
  summarise(total_export = sum(tons)) %>%
  arrange(desc(total_export))

#Joining tables
trade_partner_corn <- full_join(trade_partner_corn_export,trade_partner_corn_import)
trade_partner_corn <- trade_partner_corn %>% mutate(total_trade = total_export + total_export)

trade_partner_corn <- data.frame(trade_partner_corn)
```

```{r corn_trade_partner_plot, echo=FALSE, fig.width=14, fig.height=8}
# Plotting Results
trade_partner_corn[c(1:10),c(1,4)] %>% ggplot() +
  geom_bar(aes(x = reorder(country, -total_trade), y = total_trade, fill = reorder(country, -total_trade)), stat = "identity") +
  scale_y_continuous(labels = dollar) +
  labs(title = "Top 10 Brazil trade partners for Corn in the last 3 years",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  theme(axis.text.x = element_text(angle = 45 ),
        plot.title = element_text(hjust = 0.5, size = 16),
        legend.title = element_blank()) +
  xlab("Countries") +
  ylab("Profit(USD/Year)")
```

#### From the graph, Iran, Japan and South Korea was the main brazilian tradepartners for sugar for the last 3 years.

```{r sugar_trade_partner_summary, include=FALSE}
# Most important trade partners for brazil for sugar in the last 3 years
# Import
trade_partner_sugar_import <- bra_df %>% filter(product == "sugar" & year <= 2017 & type == "Import") %>%
  group_by(country) %>%
  summarise(total_import = sum(tons)/3) %>%
  arrange(desc(total_import))

# Export
trade_partner_sugar_export <- bra_df %>% filter(product == "sugar" & year <= 2017 & type == "Export") %>%
  group_by(country) %>%
  summarise(total_export = sum(tons)/3) %>%
  arrange(desc(total_export))


trade_partner_sugar <- full_join(trade_partner_sugar_export,trade_partner_sugar_import)
trade_partner_sugar <- trade_partner_sugar %>% mutate(total_trade = total_export + total_export)

trade_partner_sugar <- data.frame(trade_partner_sugar)
```

```{r sugar_trade_partner_plot, echo=FALSE, fig.width=14, fig.height=8}
trade_partner_sugar[c(1:10),c(1,4)] %>% ggplot() +
  geom_bar(aes(x = reorder(country, -total_trade), y = total_trade, fill = reorder(country, -total_trade)), stat = "identity") + 
  scale_y_continuous(labels = dollar) +
  labs(title = "Top 10 Brazil trade partners for Sugar in the last 3 years",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  theme(axis.text.x = element_text(angle = 45),
        plot.title = element_text(hjust = 0.5, size = 16),
        legend.title = element_blank()) +
  xlab("Countries") +
  ylab("Profit(USD/Year)")
```

#### From the graph, United Arab Emirates and Nigeria was the main brazilian trade partners for sugar for the last 3 years, followed by Bangladesh and China.

### Main states in exports for each product
#### The following analysis shows the main brazilian states in exports for each product from 1997 to 2019.

```{r state_export, include=FALSE}
#Filtering and grouping data
top_states_each_product <- bra_df %>% filter(type == "Export") %>%
  group_by(product, state) %>%
  summarise(total_tons = sum(tons), total_profit = sum(usd), mean_profit = sum(usd)/22) %>%
  arrange(product, desc(total_profit), state)

top_states_each_product <- data.frame(top_states_each_product)

# Dividing Data Set for each product
top_states_corn <- top_states_each_product %>% filter(product == "corn") %>% arrange(desc(total_profit))
top_states_wheat <- top_states_each_product %>% filter(product == "wheat") %>% arrange(desc(total_profit))
top_states_soybean_meat <- top_states_each_product %>% filter(product == "soybean_meal") %>% arrange(desc(total_profit))
top_states_soybean_oil <- top_states_each_product %>% filter(product == "soybean_oil") %>% arrange(desc(total_profit))
top_states_soybeans <- top_states_each_product %>% filter(product == "soybeans") %>% arrange(desc(total_profit))
top_states_sugar <- top_states_each_product %>% filter(product == "sugar") %>% arrange(desc(total_profit))
```

```{r corn_state_export, echo=FALSE, fig.width=14, fig.height=8}
top_states_corn %>% ggplot() +
  geom_bar(aes(x = reorder(state, -mean_profit), y = total_profit, fill = reorder(state, -mean_profit)), position = "dodge", stat = "identity") +
  scale_y_continuous(trans = log10_trans(),
                     labels = dollar) +
  labs(title = "Ranking of States for Corn exports",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(hjust = 0.5, size = 16),
        legend.title = element_blank()) +
  xlab("States") +
  ylab("Profit(USD/Year)")
```

#### The main states for corn exports are: MT, PR, GO, MS and SP.

```{r wheat_state_export, echo=FALSE, fig.width=14, fig.height=8}
top_states_wheat %>% ggplot() +
  geom_bar(aes(x = reorder(state, -mean_profit), y = total_profit, fill = reorder(state, -mean_profit)), position = "dodge", stat = "identity") +
  scale_y_continuous(trans = log10_trans(),
                     labels = dollar) +
  labs(title = "Ranking of States for Wheat exports",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(hjust = 0.5, size = 16),
        legend.title = element_blank()) +
  xlab("States") +
  ylab("Profit(USD/Year)")
```

#### The main states for Wheat exports are: RS, PR, SC, SP and MS.

```{r soybean_meat_state_export, echo=FALSE, fig.width=14, fig.height=8}
top_states_soybean_meat %>% ggplot() +
  geom_bar(aes(x = reorder(state, -mean_profit), y = total_profit, fill = reorder(state, -mean_profit)), position = "dodge", stat = "identity") +
  scale_y_continuous(trans = log10_trans(),
                     labels = dollar) +
  labs(title = "Ranking of States for Soybean Meat exports",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(hjust = 0.5, size = 16),
        legend.title = element_blank()) +
  xlab("States") +
  ylab("Profit(USD/Year)")
```

#### The main states for Soybean Meat exports are: MT, PR, RS, GO and BA.

```{r soybean_oil_state_export, echo=FALSE, fig.width=14, fig.height=8}
top_states_soybean_oil %>% ggplot() +
  geom_bar(aes(x = reorder(state, -mean_profit), y = total_profit, fill = reorder(state, -mean_profit)), position = "dodge", stat = "identity") +
  scale_y_continuous(trans = log10_trans(),
                     labels = dollar) +
  labs(title = "Ranking of States for Soybean Oil exports",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(hjust = 0.5, size = 16),
        legend.title = element_blank()) +
  xlab("States") +
  ylab("Profit(USD/Year)")
```

#### The main states for Soybean Oil exports are: PR, RS, MT, SC and GO.

```{r soybeans_state_export, echo=FALSE, fig.width=14, fig.height=8}
top_states_soybeans %>% ggplot() +
  geom_bar(aes(x = reorder(state, -mean_profit), y = total_profit, fill = reorder(state, -mean_profit)), position = "dodge", stat = "identity") +
  scale_y_continuous(trans = log10_trans(),
                     labels = dollar) +
  labs(title = "Ranking of States for Soybeans exports",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(hjust = 0.5, size = 16),
        legend.title = element_blank()) +
  xlab("States") +
  ylab("Profit(USD/Year)")
```

#### The main states for Soybeans exports are: MT, PR, RS, GO and MS.

```{r sugar_state_export, echo=FALSE, fig.width=14, fig.height=8}
top_states_sugar %>% ggplot() +
  geom_bar(aes(x = reorder(state, -mean_profit), y = total_profit, fill = reorder(state, -mean_profit)), position = "dodge", stat = "identity") +
  scale_y_continuous(trans = log10_trans(),
                     labels = dollar) +
  labs(title = "Ranking of States for Sugar exports",
       caption = "Data from: http://comexstat.mdic.gov.br/") +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(hjust = 0.5, size = 16),
        legend.title = element_blank()) +
  xlab("States") +
  ylab("Profit(USD/Year)")
```

#### The main states for Sugar exports are: SP, MG, PR, AL and MS.

### Export prediction

#### This last study forecast the export for corn soybeans and soybean_meal from 2019 to 2020.

```{r forecast, include=FALSE}
# Modelling
# Forecast for Soybeans, Soybean Meal and Corn exports in tons for the next 11 years (2020, 2030)
product_ton_year  <- bra_df %>% filter(type == "Export" & (product == "corn" | product == "soybeans" | product == "soybean_meal")) %>%
  group_by(product, year) %>%
  summarise(total_tons = sum(tons))

product_ton_year <- data.frame(product_ton_year)

corn_ton_year <- product_ton_year %>% filter(product == "corn")
soybeans_ton_year <- product_ton_year %>% filter(product == "soybeans")
soybean_meal_ton_year <- product_ton_year %>% filter(product == "soybean_meal")

corn_ton_year$product <- NULL
soybeans_ton_year$product <- NULL
soybean_meal_ton_year$product <- NULL

# Time Series Converting
corn_TSobj <- ts(corn_ton_year)
soybeans_TSobj <- ts(soybeans_ton_year)
soybean_meal_TSobj <- ts(soybean_meal_ton_year)

# TBATS autocorrelations descriptions
# T: Trigonometric terms for seasonality
# B: Box-Cox transformations for heterogeneity
# A: ARMA errors for short-term dynamics
# T: Trend
# S: Seasonal (including multiple and non-integer periods)
tmp_tbats_corn <- tbats(corn_TSobj[,2])
tmp_tbats_soybeans <- tbats(soybeans_TSobj[,2])
tmp_tbats_soybeans_meal <- tbats(soybean_meal_TSobj[,2])

# Forecast by autocorrelations
tbats_corn <- forecast::forecast(tmp_tbats_corn)
tbats_soybeans <- forecast::forecast(tmp_tbats_soybeans)
tbats_soybean_meal <- forecast::forecast(tmp_tbats_soybeans_meal)

# Apply forecast Method for each time series
tbats_corn <- holt(corn_TSobj[,2], h = 11)
tbats_soybeans <- holt(soybeans_TSobj[,2], h = 11)
tbats_soybean_meal <- holt(soybean_meal_TSobj[,2], h = 11)

# Create Dataframe for forecasted values
forecast_corn <- data.frame(c(2020:2030), c(tbats_corn$mean))
forecast_soybeans <- data.frame(c(2020:2030), c(tbats_soybeans$mean))
forecast_soybean_meal <- data.frame(c(2020:2030), c(tbats_soybean_meal$mean))

## Forecasted Data Frame
colnames(forecast_corn) <- c("year", "total_tons")
colnames(forecast_soybeans) <- c("year", "total_tons")
colnames(forecast_soybean_meal) <- c("year", "total_tons")

# Joining tables
corn_ton_year_forecast <- bind_rows(corn_ton_year, forecast_corn)
soybeans_ton_year_forecast <- bind_rows(soybeans_ton_year, forecast_soybeans) 
soybean_meal_ton_year_forecast <- bind_rows(soybean_meal_ton_year, forecast_soybean_meal)
```

```{r forecast_plots, echo=FALSE, fig.width=14, fig.height=8}
#Plotting Results
corn_ton_year_forecast %>% ggplot +
  geom_point(aes(x = year, y = total_tons), lwd = 1.1) +
  geom_vline(xintercept = 2019) +
  theme(plot.title = element_text(hjust = 0.5, size = 16)) +
  ggtitle("Total Corn production Prediction") +
  ylab("Quantity(ton)") + 
  xlab("Years")

#Plotting Results
soybeans_ton_year_forecast %>% ggplot +
  geom_point(aes(x = year, y = total_tons), lwd = 1.1) +
  geom_vline(xintercept = 2019) +
  theme(plot.title = element_text(hjust = 0.5, size = 16)) +
  ggtitle("Total Soybeans production Prediction") +
  ylab("Quantity(ton)") + 
  xlab("Years")

#Plotting Results
soybean_meal_ton_year_forecast %>% ggplot +
  geom_point(aes(x = year, y = total_tons), lwd = 1.1) +
  geom_vline(xintercept = 2019) +
  theme(plot.title = element_text(hjust = 0.5, size = 16)) +
  ggtitle("Total Soybean Meal production Prediction") +
  ylab("Quantity(ton)") + 
  xlab("Years")
```